Hunting experience shapes individual foraging specialization and predator-prey interactions in an online videogame

### 2023-05-09 manuscript/Chapter2-manuscript.Rmd
finished updating the results

Out of all five models, the two that best predicted the data, with an almost identical expected log pointwise density (elpd), were the models that controlled for the prey group's speed (Table 1). Predators differed substantially in the relationship between hunting success and cumulative experience (i.e. expertise: Figure 1A-B). For instance, the success of some predators steadily increased with experience (Figure 1B, Appendix S1: Figure S1A) while others showed a decrease in success with experience (Figure 1B, Appendix S1: Figure S1B). Overall, `r toString(scales::percent(191/253))` of the population displayed an >10% increase in hunting success, `r toString(scales::percent(37/253))` displayed a >10% decrease in hunting success, and the remaining `r toString(scales::percent(25/253))` had a lesser change in success ranging between -10% and +10%.

We predicted that the relationship between hunting success and experience would vary among predators because they should not encounter the same prey. After controlling for the mean speed of the prey group within a match, we did not find evidence confirming this prediction, because differences among predators remained important (Figure 1B). However, many individual curves appear to have shifted upwards in the model where we control for prey speed (Figure 1B) vs in the model where we do not control for prey speed (Figure 1A). Moreover, in the models where we did not account for prey speed, the global trend of predator experience was nearly flat (results not shown). Yet, we see that there is an increase up to 13% in the probability of capturing all prey with experience in the population when we account for prey speed (Figure 1C).

## Effect of experience on the development of behavioural specialization

The mean speed of the predator population did not change with experience (i.e. intercept of predator speed (mean); Figure 2 and Appendix S1: Table S1). There was also strong evidence that foraging specialization in the population remained stable across experience levels (i.e. intercept of predator speed (sigma); Figure 2 and Appendix S1: Table S1).

Individual predators differed slightly in their average movement speed, but differences did not increase at higher experience levels (i.e. predator ID (mean); Figure 2 and Appendix S1: Table S2). However, individuals displayed differences in their foraging specialization, with a 0.1 unit increase in standard deviation from novice to advanced (i.e. predator ID (sigma) is higher at the advanced level; Figure 2 and Appendix S1: Table S2). Thus, as individual predators gain experience, the population becomes slightly more heterogeneous along the specialist-flexible hunter continuum. Moreover, some individual predators switched from a flexible to a specialized hunting tactic, and vice-versa, as they gained experience (Figure 3). Importantly, 12.6% of the predator population displayed a large change in foraging specialization/flexibility with experience (i.e. $\ge$0.2 change in standard deviation), 44.7% displayed lower changes (i.e. >0.05 and <0.2 change in standard deviation), while 42.7% remained relatively stable (i.e. $\le$0.05 change in standard deviation).

Contrary to predictions, there was strong evidence that predators did not differ in the speed of the prey groups that they encountered on average (i.e. predator ID (mean); Appendix S1: Table S2). Moreover, differences in standard deviation values across experience were negligible (i.e. predator ID (mean); Figure 2). However, individuals differed in the heterogeneity of the groups of prey encountered across the study period (i.e. predator ID (sigma); Appendix S1: Table S2). Hence, some predators encountered prey groups with similar average speeds, while others encountered prey groups with varying average speeds throughout the study period. The differences in the prey encountered among predators increased only slightly with experience (i.e. predator ID (sigma); Figure 2).

@@

*continuer ici les modifications*

At all experience levels, predators that were on average faster (mean) also had a lower intra-individual variability (IIV; Figure 3 and Appendix S2: Table S1). Thus, faster predators specialized on fast movement, while slower predators were more flexible in their movement through time. Predators that were on average faster (mean) also encountered groups of prey that were on average faster (mean) (Figure 3 and Appendix S2: Table S1). At the three experience levels, there were only weak correlations between IIV in predator speed and mean prey speed (Figure 3 and Appendix S2: Table S1). However, predators that encountered groups of prey with similar speeds (IIV) specialized in their hunting tactic (IIV) (Figure 3 and Appendix S2: Table S1). Hence, predators that encountered groups of prey with similar average speeds were more specialized, while those that encountered groups of prey with different average speeds were more flexible. For all the results described above, the correlations strengthen as predators gain experience (Figure 3 and Appendix S2: Table S1).

@@

We did not detect strong correlations between the predators' mean speed and their hunting success (Figure 3 and Appendix S2: Table S1), nor between the predators' IIV in speed and their hunting success (Figure 3 and Appendix S2: Table S1), suggesting that slower/flexible and faster/specialist predators were both equally successful. However, novice individuals that encountered groups of prey with faster speeds were less successful (Figure 3 and Appendix S2: Table S1). This relationship was decoupled at the later experience levels, such that encounters with faster prey at intermediate and advanced stages were not as strongly correlated with hunting success anymore (Figure 3 and Appendix S2: Table S1). Thus, the consequences of hunting faster prey were reduced at greater experience levels.

@@

@@

**Figure 2.** Posterior median differences and credible intervals between each predator experience level's parameters. Each panel shows the experience levels that are compared. The parameter compared is on the y axis, and the difference of the values predicted by the MDHGLM is on the x axis. The intercepts correspond to mean population speed and specialization. The predator ID on the mean are individual differences in mean speed, mean speed of prey groups encountered, and success. The predator ID on the dispersion (i.e. sigma) are individual differences in specialization, and in the heterogeneity of prey encountered.

@@

**Figure 4.** Correlation matrices of the MDHGLM relating the mean and intra individual variability (IIV) in speed of the predators, the mean and IIV in speed of the prey they encountered, and mean hunting success. Larger dark blue circles indicate stronger positive correlations (positive values on the color legend), while larger dark red circles indicate stronger negative correlations (negative values on the color legend). A) Correlations when predators were novice hunters. B) Correlations when predators were advanced hunters. C) Posterior median differences and credible intervals comparing the correlations between traits when predators were advanced vs novice. The median difference is displayed on the x axis and the trait correlations being compared are displayed on the y axis.

>>>>

Out of all five models, the two that best predicted the data, with very similar expected log pointwise densities (elpd), were the models that controlled for the prey group's speed (Table 1). Predators differed substantially in the relationship between hunting success and cumulative experience (i.e. expertise: Figure 1A-B). For instance, the success of some predators steadily increased with experience (Figure 1B, Appendix S1: Figure S1A) while others showed a decrease in success with experience (Figure 1B, Appendix S1: Figure S1B). Overall, `r toString(scales::percent(191/253))` of the population displayed an >10% increase in hunting success, `r toString(scales::percent(37/253))` displayed a >10% decrease in hunting success, and the remaining `r toString(scales::percent(25/253))` had a lesser change in success ranging between -10% and +10%.

We did not find evidence that differences among predators in the relationship between hunting success and experience was mediated by prey encounters because differences remained high even after controlling for the mean speed of the prey groups (Figure 1B). However, many individual curves appear to have shifted upwards after accounting for prey speed (Figure 1B) vs when we do not control for prey speed (Figure 1A). Moreover, in the models where we did not account for prey speed, the global trend of predator experience was nearly flat (results not shown). Yet, we see that there is an increase up to 13% in the probability of capturing all prey with experience in the population when we account for prey speed (Figure 1C).

## Effect of experience on behavioural specialization and hunting success

The mean speed of the predator population did not change with experience (i.e. intercept (mean) of predator speed; Figure 2 and Appendix S1: Table S1). There was also strong evidence that foraging specialization in the population remained stable across experience levels (i.e. intercept (sigma) of predator speed; Figure 2 and Appendix S1: Table S1). We did not find evidence that the predator population had a change in mean success with experience (intercept (mean) of hunting success; Figure 2 and Appendix S1: Table S1). There was, however, a decrease in among individual differences in success at the intermediate stage, which was maintained at the advanced stage (predator ID (mean) of hunting success; Figure 2 and Appendix S1: Table S2).

Individual predators differed slightly in their average movement speed across experience levels (i.e. predator ID (mean); Figure 2 and Appendix S1: Table S2). However, there was strong evidence that they displayed differences in foraging specialization, with a 0.1 unit increase in standard deviation from novice to advanced, indicating that differences increased slightly (i.e. predator ID (sigma) is higher at the advanced level; Figure 2 and Appendix S1: Table S2). Moreover, some individual predators switched from a flexible to a specialized hunting tactic, and vice-versa, as they gained experience (Figure 3). Importantly, 12.6% of the predator population displayed a large change in foraging specialization/flexibility with experience (i.e. $\ge$0.2 change in standard deviation), 44.7% displayed lower changes (i.e. >0.05 and <0.2 change in standard deviation), while 42.7% remained relatively stable (i.e. $\le$0.05 change in standard deviation).

Contrary to predictions, there was strong evidence that predators did not differ in the speed of the prey groups that they encountered on average across experience (i.e. predator ID (mean); Figure 2 and Appendix S1: Table S2). However, individuals differed in the heterogeneity of the groups of prey encountered across the study period (i.e. predator ID (sigma); Appendix S1: Table S2). Hence, some predators encountered prey groups with similar average speeds, while others encountered prey groups with varying average speeds throughout the study period. The differences in the prey encountered among predators increased only slightly with experience (i.e. predator ID (sigma); Figure 2).

@@

Predators that were on average faster (mean speed) had a lower intra-individual variability (IIV speed) (Figure 4A-B and Appendix S4: Table S1). Thus, faster predators specialized on fast movement, while slower predators were more flexible in their movement through time. As predators gained experience, these strategies were increasingly defined as there was a 0.16 unit increase in the correlation at the advanced stage (Figure 4C). Predators that were on average faster (mean speed) also encountered groups of prey that were on average faster (mean prey speed) and more similar (IIV prey speed) (Figure 4A-B and Appendix S4: Table S1). There was however only weak to no evidence that these correlations increased with experience (Figure 4C)

There was almost no correlation between IIV in predator speed and mean prey speed when predators were novice (Figure 4A). There was however a 0.17 unit increase in this correlation at the advanced stage, such that specialist hunters at this stage tended to encounter faster prey (Figure 4C). The predator IIV in speed and IIV in prey speed were moderately positively correlated (Figure 4A-B and Appendix S4: Table S1). Hence, predators that encountered groups of prey with similar speeds were more specialized in their tactic, and those that encountered groups of prey with varying speeds were more flexible. We did not find evidence that this correlation increased with experience (Figure 4C).

@@

The predator's mean speed and IIV were only weakly negatively correlated with hunting success across experience (Figure 4 and Appendix S4: Table S1), suggesting that slower/flexible and faster/specialist predators were both equally successful. However, there was strong evidence that individuals who encountered groups of prey with faster speeds were less successful (Figure 4A-B and Appendix S4: Table S1). Predators also had greater success across the study when encountering more variable groups of prey (Figure 4A-B and Appendix S4: Table S1). Both relationships remained structurally stable across experience (Figure 4-C)

@@

*simplifier la 1ere phrase*

@@

**Figure 2.** Posterior median differences and highest posterior density intervals between each predator experience level's parameters. Each panel shows the experience levels that are compared. The parameter compared is on the y axis, and the difference of the values predicted by the MDHGLM is on the x axis. The intercepts correspond to mean population speed and specialization. The predator ID on the mean are individual differences in mean speed, mean speed of prey groups encountered, and success. The predator ID on the dispersion (i.e. sigma) are individual differences in specialization, and in the heterogeneity of prey encountered.

@@

**Figure 4.** Correlation matrices of the MDHGLM relating the mean and intra individual variability (IIV) in speed of the predators, the mean and IIV in speed of the prey they encountered, and mean hunting success. Larger dark blue circles indicate stronger positive correlations (positive values on the color legend), while larger dark red circles indicate stronger negative correlations (negative values on the color legend). A) Correlations when predators were novice hunters. B) Correlations when predators were advanced hunters. C) Posterior median differences and highest posterior density intervals comparing the correlations between traits when predators were advanced vs novice. The differences were computed on the absolute values of the correlations since the signs did not change with experience, such that positive values indicate an increase in correlation at the advanced stage. The median difference is displayed on the y axis and the trait correlations being compared are displayed on the x axis.
